Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem

### 2015-12-07 acta_manuscript.tex
added tikz figures

% introduce \Poincare section theory and math
\Poincare maps are a useful tool in the analysis of the flow near periodic orbits in the three-body problem.
We let \( \Sigma \) be a hypersurface of section chosen such that all trajectories in the vicinity of a state \( q \in \Sigma \) cross \( \Sigma \) transversely and in the same direction.
A \Poincare map, \( P(q) = \phi(T;q) \), maps the state of a trajectory from one intersection to the next.
Choosing a section in this manner results in a \Poincare section as shown in~\cref{fig:poincare_map}.
This allows for greater insight into the stability and dynamics of periodic solutions of dynamic system as a fixed point on the \Poincare section corresponds to a periodic orbit while movement on the section is associate with the stability of neighboring trajectories. 
For example, \Poincare maps have been used to prove the existence of homoclinic orbits, orbits which are both forward and backward asymptotic to a single unstable periodic orbit, and heteroclinic orbits, which join different periodic orbits~\cite{conley1968,koon2000b}.
These dynamic features have been shown to play a vital role in the movement of natural bodies as well as critical for spacecraft missions~\cite{lo1997,gomez2001}.
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{poincare_map}
	\caption{\Poincare map\label{fig:poincare_map}}
\end{figure}

@@

In this paper, we seek to approximate the reachability set on a \Poincare set by solving a related optimal control problem. 
We choose our \Poincare section in a similar manner to those used previously for the design of transfers via invariant manifolds.
The \Poincare section is chosen to intersect transversally with trajectories emanating from the initial orbit. 
In the case of a periodic orbit the trajectories will cross the \Poincare section at two distinct fixed points every half period.
The addition of low thrust propulsion allows us to enlarge the set of trajectories achievable. 
\Cref{fig:reachability_set} illustrates how, without any control input, trajectories will intersect with the \Poincare section at \( \vec{x}_n \). 
However, the addition of low thrust propulsion allows the spacecraft to depart from the natural dynamics and intersect the \Poincare section at a different location.
We use a cost function to define a distance metric on the \Poincare section from the control-free intersection to an intersection under the influence of the control input.
Maximization of this distance enables us to generate the largest reachability set under the bounded control input.
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{reachability_set}
	\caption{Reachability set\label{fig:reachability_set}}
\end{figure}

>>>>

% introduce \Poincare section theory and math
\Poincare maps are a useful tool in the analysis of the flow near periodic orbits in the three-body problem.
We let \( \Sigma \) be a hypersurface of section chosen such that all trajectories in the vicinity of a state \( q \in \Sigma \) cross \( \Sigma \) transversely and in the same direction.
A \Poincare map, \( P(q) = \phi(T;q) \), maps the state of a trajectory from one intersection to the next.
Choosing a section in this manner results in a \Poincare section as shown in~\cref{fig:poincare_map}.
This allows for greater insight into the stability and dynamics of periodic solutions of dynamic system as a fixed point on the \Poincare section corresponds to a periodic orbit while movement on the section is associate with the stability of neighboring trajectories. 
For example, \Poincare maps have been used to prove the existence of homoclinic orbits, orbits which are both forward and backward asymptotic to a single unstable periodic orbit, and heteroclinic orbits, which join different periodic orbits~\cite{conley1968,koon2000b}.
These dynamic features have been shown to play a vital role in the movement of natural bodies as well as critical for spacecraft missions~\cite{lo1997,gomez2001}.
\begin{figure}
	\centering
	\begin{scaletikzpicturetowidth}{0.5\textwidth}
	\begin{tikzpicture}[scale=\tikzscale]
		\coordinate [label=left:\textcolor{black}{\LARGE \(\vec{q}_0\)}] (x0) at (0,0);
		\coordinate [label=right:\textcolor{black}{\LARGE  \(\vec{q}_k\)}] (xn) at (-1,-1);
		\coordinate [label=left:\textcolor{black}{\LARGE  \(\Sigma\)}] (sigma) at (-4,3);
		\coordinate [label=below:\textcolor{black}{\LARGE  \(P(\vec{q})\)}] (P) at (0,-3.5);
		% define the path of the flow with coordinates
		\coordinate [label=right:\textcolor{blue}{}] (f1) at (5,-2);
		\coordinate [label=below:\textcolor{black}{\LARGE  \(\phi(t,\vec{q})\)}] (f2) at (2,-5);
		\coordinate [label=right:\textcolor{blue}{}] (f3) at (-4,-4);
		\coordinate [label=right:\textcolor{blue}{}] (f4) at (-4,-1);

	%	\draw[help lines] (-10,-10) grid (10,10); %grid
		\filldraw [black] (x0) circle [radius=3pt];
		\filldraw [black] (xn) circle [radius=3pt];

		\draw [ultra thick,black,->-](x0) to[out=20,in=90,distance=2cm] (f1) to[out=-90,in=0,distance=2cm] (f2) to[out=180,in=-45,distance=2cm] (f3) to[out=135,in=-135,distance=2cm] (f4) ;
		\draw [ultra thick, black,dashed] (f4) to[out=45,in=120,distance=1cm] (xn);

		\draw [ultra thick] plot [smooth cycle, tension=0.1, rotate=5] coordinates { (-4,-3) (4,-3) (4,3) (-4,3) };
	\end{tikzpicture}
	\end{scaletikzpicturetowidth}
	\caption{Diagram of the \Poincare map\label{fig:poincare_map}}
\end{figure}

@@

In this paper, we seek to approximate the reachability set on a \Poincare set by solving a related optimal control problem. 
We choose our \Poincare section in a similar manner to those used previously for the design of transfers via invariant manifolds.
The \Poincare section is chosen to intersect transversally with trajectories emanating from the initial orbit. 
In the case of a periodic orbit the trajectories will cross the \Poincare section at two distinct fixed points every half period.
The addition of low thrust propulsion allows us to enlarge the set of trajectories achievable. 
\Cref{fig:reachability_set} illustrates how, without any control input, trajectories will intersect with the \Poincare section at \( \vec{x}_n \). 
However, the addition of low thrust propulsion allows the spacecraft to depart from the natural dynamics and intersect the \Poincare section at a different location.
We use a cost function to define a distance metric on the \Poincare section from the control-free intersection to an intersection under the influence of the control input.
Maximization of this distance enables us to generate the largest reachability set under the bounded control input.
\begin{figure}
	\centering
%	\includegraphics[width=0.5\textwidth]{reachability_set_hand}
	\begin{scaletikzpicturetowidth}{0.5\textwidth}
	\begin{tikzpicture}[scale=\tikzscale]
		\coordinate [label=left:\textcolor{black}{\LARGE \(\vec{x}_0\)}] (x0) at (-1,-2);
		\coordinate [label=below:\textcolor{black}{\LARGE  \(\vec{x}_n\)}] (xn) at (1,1);
		\coordinate [label=left:\textcolor{black}{\LARGE  \(\Sigma\)}] (sigma) at (-4,3);
		\coordinate [label=below:\textcolor{black}{\LARGE  \(P(\vec{x})\)}] (P) at (0,-3.5);
		% define the path of the flow with coordinates
		\coordinate [label=right:\textcolor{black}{}] (f1) at (5,-2);
		\coordinate [label=below:\textcolor{black}{\LARGE  \(\phi(t,\vec{x}_0)\)}] (f2) at (2,-5);
		\coordinate [label=right:\textcolor{black}{}] (f3) at (-4,-4);
		\coordinate [label=right:\textcolor{black}{}] (f4) at (-4,-1);

	%	\draw[help lines] (-10,-10) grid (10,10); %grid
		\filldraw [black] (x0) circle [radius=3pt];
		\filldraw [black] (xn) circle [radius=3pt];

		\draw [ultra thick,black,->-](x0) to[out=20,in=90,distance=2cm] (f1) to[out=-90,in=0,distance=2cm] (f2) to[out=180,in=-45,distance=2cm] (f3) to[out=135,in=-135,distance=2cm] (f4) ;
		\draw [ultra thick, black,dashed,->] (f4) to[out=45,in=180,distance=1cm] ($(xn)-(2,0)$);

		\draw [ultra thick] plot [smooth cycle, tension=0.1, rotate=5] coordinates { (-4,-3) (4,-3) (4,3) (-4,3) };

		\draw [thick,dashed] (xn) circle [radius=2cm]; % reachability set

		\draw [thick,->] (xn) -- ($(xn) + (2.5,0)$);
		\draw [thick,rotate=45,->] (xn) -- ($(xn) + (2.5,0)$);
		\draw ($(xn) + (1,0)$) arc [start angle=0,end angle=45, radius=1];
		\node [draw=none] at (2.2,1.5) {\Large \(\theta\)};
		\draw [decorate,decoration={brace,amplitude=5pt},rotate=45] (xn) -- ($(xn) + (2,0)$);
		\node [draw=none] at ($ (xn) + (0,1) $) {\Large \( J \)};
	\end{tikzpicture}
	\end{scaletikzpicturetowidth}
	\caption{Reachability set on a \Poincare section\label{fig:reachability_set}}
\end{figure}
